---
title: Weather Conditions and their Influence on human Mobility Patterns
subtitle: A descriptive approach
author: Lars Weidinger & Javier Feller
format:
  html:
    code-fold: true
execute:
  warning: false
  message: false
lang: en
bibliography: bibliography.bib
---

```{r preprocessing}
#| code-summary: preprocessing
source("preprocessing.R") # this runs the whole preprocessing file
``` 

# Notes from Project Proposal Call
Meteo Daten Stunde vs. Tag --> ausprobieren

Sample zu klein wahrscheinlich, Stunde sieht man probably mehr Muster
andere Studierende? 

Map Matching --> nicht nötig für unser Projekt

Timeline vs. ArcGIS Earth --> wie kann man das vergleichen?
Filter machen für die Abschnitte, bei denen wir die gleichen Informationen haben
--> falls zu aufwändig, dann treffen wir eine pragmatische Entscheidung und halten das so fest (Vereinfachung) 

1) Anschauen, ob man mit Meteodaten (1 Tag) Unterschiede auf dem gesamten Datensatz sieht. 
- Zu Beginn explorativ anschauen, ob es Unterschiede gibt von Wochentagen und Wochenenden 

2) Dann genauer schauen mit Meteodaten (1h oder 15min) 

3) Dann noch genauer schauen und zwar nur dort, wo man gelaufen ist.
- gibt es Unterschiede in Laufdistanz oder Laufgeschwindigkeit je nach Temperatur?

4) Falls man keine krassen Unterschiede sieht --> zwei Ansätze miteinander vergleichen (z.B. 15min vs. 1h oder 1 Tag) 
--> geht um konzeptueller Approach, Ergebnisse sind nicht so relevant

Alternative Datensätze: Tageslicht? Stürme? 


# Abstract

# Introduction

@laube2011 investigate how temporal scale affects the calculation of movement parameters (speed, sinuosity and turning angle) of animal trajectories. 

# Material and Methods

## Load Packages
```{r}
library(pacman)

# Load and if not found in the library will attempt to install the package
pacman::p_load(here,
               sf,
               readr,
               tmap,
               dplyr,
               ggplot2,
               tidyr,
               jsonlite,
               lubridate, # temporal import from JSON file
               tidyverse,
               leaflet, # interactive maps
               tmap, # interactive maps
               concaveman, # construct concave hulls
               install = TRUE)

```

## Set CRS
```{r}
crs_lv95  <- 2056
crs_wgs84 <- 4326
```

## Import Data
See the preprocessing.R file where we import:  
- weather data  
- tracking data  
- boundary data?

```{r}
stadtgrenze <- read_sf("data/Gemeindegrenzen_OGD/Gemeindegrenzen_-OGD.gpkg", layer = "up_gemeinden_f") |> # take the polygon
  filter(gemeindename == "Zürich") |> # of the city of Zurich
  select(geom)
```

## Visualise Weather Data Daily

### Locating Weather Stations
Since the data did not provide coordinates of the measuring stations, we manually took the CH1903+ coordinates of points that have to be close to the actual measuring stations using map.geo.admin.ch. We will ask the City of Zurich for the exact locations of the stations but for now this is our approximation for the moment. 

```{r}
weather_stations <- data.frame(
  Standort = c(
    "Zch_Heubeeribüel",
    "Zch_Rosengartenstrasse",
    "Zch_Schimmelstrasse",
    "Zch_Stampfenbachstrasse"
  ),
  E = c(
    2685280.12,
    2682050.59,
    2682105.79,
    2683454.71
  ),
  N = c(
    1248962.89,
    1249880.93,
    1247104.63,
    1248019.62
  )
)

weather_stations_sf <- st_as_sf(
  weather_stations,
  coords = c("E", "N"),
  crs = crs_lv95)
```

```{r}
ggplot() +
  geom_sf(data = stadtgrenze, fill = "#D5F4FF", color = "black") +
  geom_sf(data = weather_stations_sf, aes(color = Standort), size = 3) +
  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Weather Stations within Zurich",
    color = "Station",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Constrcut Voronoi Polygons around Measuring Stations
Since we need to assign weather data to our stops and later our segments we need to relate the two datasets spatially. For that, we construct Voronoi polygons around the measuring stations and assign each tracking point the weather values from the station it is closest to (= inside the Voronoi polygon).

```{r}
# both datasets are in CH1903+

# create a union of points
pts_union <- st_union(weather_stations_sf)

# create Voronoi tessellation with an envelope
# st_voronoi has an envelope argument. Instead of generating a huge (or to small) tessellation we can set the scope here
envelope_stadtgrenze <- st_as_sfc(st_bbox(stadtgrenze))

voronoi <- st_voronoi(
  pts_union,
  envelope = envelope_stadtgrenze)

# extract polygons
voronoi_polygons <- st_collection_extract(voronoi, "POLYGON")

# convert to sf object
voronoi_sf <- st_sf(
  Standort = weather_stations_sf$Standort,
  geometry = st_cast(voronoi_polygons))

# clip to city boundary
voronoi_clipped <- sf::st_intersection(voronoi_sf, stadtgrenze)
```

```{r}
ggplot() +
  geom_sf(data = voronoi_clipped, aes(fill = Standort)) +
  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Voronoi Polygons around\n Weather Stations in Zurich",
    color = "Station",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Assign Google Timeline Data to Voronoi Polygons
We can now use a spatial join to assign the corresponding weather station to each one of the Google Timeline tracking points.  
Since we traveled quite a lot outside of the city of Zurich, we omit the data outside the city (NA values), which is approximately 60% of the data points. 

```{r}
# both datasets are in CH1903+

# spatial join: assign polygon attributes (Standort) to points
google_timeline_sf_lv95 <- st_join(
  google_timeline_sf_lv95,
  voronoi_clipped["Standort"],
  join = st_within) |>
  
  drop_na() # omit NA values (quite a lot of them)
```

### Assign Weather Data to Google Timeline Data
Now that we have assigned each data point to its closest weather station, we need to assign them their relevant weather values. We decided on using temperature (T) and rain (RainDur). One problem is, that the station Heubeeribüel does not collect RainDur information. 

```{r}
# based on the exercises of week 5, Task 3 ff we first create a join key so that we can later match them exactly

google_timeline_sf_lv95 <- google_timeline_sf_lv95 |>
  mutate(
    Date = as.Date(time)) # since we only have one measurement per night

weather_daily_df <- weather_daily_df |>
  mutate(
    Date = as.Date(Datum)) # since we only have one measurement per night
```

```{r}
# prepare weather data
weather_daily_df_wide <- weather_daily_df |>
  filter(Parameter %in% c("T", "RainDur")) |> # we only want T and RainDur
  select(Date, Standort, Parameter, Wert) |>
  pivot_wider(                               # we need the parameters as columns
    names_from = Parameter,
    values_from = Wert
  ) |>
  rename(
    Temperature = T,
    RainDur = RainDur
  )
```

```{r}
# join the two datasets
google_timeline_weather <- google_timeline_sf_lv95 |>
  left_join(
    weather_daily_df_wide,
    by = c("Date", "Standort")
  )
```


## Visualise Activity Space

To answer RQ II we need to construct some sort of measurement for activity space. We decided on constructing concave hulls around the trips for each day. We then can calculate the area of these hulls, take the average weather values across the whole hull and analyse if the area is larger when T is higher or P is lower.

We decided to use concave hulls rather than convex hulls since they are less prone to outliers. Even though there are some convex hull implementations that allow for some sort of variable to be manipulated in order to exclude outliers (e.g. adehabitatHR::mcp()), the concave hull is more suited for our use case. We could have also used Alpha shapes (alphahull::) which accounts for holes in our data but this is not needed since we are dealing with cotinuous travel data. 

QUELLE: GEO 872, 3. Vlsg 02.10.24 Point Patterns - Polygon Delineation

Alternative: Standard devitational ellipse (SDE): forms a closed shape that represents distributional properties of the point set but does not enclose all points.

### Construct Concave Hulls

In the first approach we called concaveman() on the entire sfc column (all geometries at once), not on each row individually. It computed one hull over all points and that same result gets recycled into every row which resulted in all the hulls being the same. We asked Claude AI (Model Sonnet 4.6) for help and it delivered the following solution using lapply.

```{r}
# Step 1: union points by date
google_timeline_hulls <- google_timeline_sf_lv95 |>
  group_by(Date) |>
  summarise(geometry = st_union(geometry), .groups = "drop") |>
  st_as_sf()

# Step 2: build concave hull for each date
hulls_list <- lapply(google_timeline_hulls$geometry, function(g) {
  coords <- st_coordinates(g)[, c("X", "Y")]
  
  # need at least 3 unique points for a polygon
  if (is.null(nrow(coords)) || nrow(unique(coords)) < 3) return(NA)
  
  st_polygon(list(concaveman(coords)))
})

# Step 3: check which dates failed
failed_dates <- google_timeline_hulls$Date[sapply(hulls_list, function(x) identical(x, NA))]
print(failed_dates)  # inspect which dates are problematic

# Step 4: remove failed dates and assign geometry
google_timeline_hulls <- google_timeline_hulls[!sapply(hulls_list, function(x) identical(x, NA)), ]

hulls_list <- hulls_list[!sapply(hulls_list, function(x) identical(x, NA))]

# Step 5: assign hulls back as sfc column
google_timeline_hulls$geometry <- st_sfc(hulls_list, crs = st_crs(google_timeline_sf_lv95))
```

```{r}
# calculate area
google_timeline_hulls <- google_timeline_hulls |>
  mutate(
    area_m2 = as.numeric(st_area(geometry)),
    area_km2 = area_m2 / 1000000)
```

### Assign Weather Values
```{r}
# average weather values across whole hull
daily_weather <- google_timeline_weather |>
  st_drop_geometry() |>
  group_by(Date) |>
  summarise(
    Temperature = mean(Temperature, na.rm = TRUE),
    RainDur = mean(RainDur, na.rm = TRUE)
  )

daily_hulls_weather <- google_timeline_hulls |>
  left_join(daily_weather, by = "Date")
```

### Statistical Analysis
We are under the assumption that the the area is bigger when temperature is higher or RainDur is higher.

```{r}
# visualise the variable Temperature
ggplot(daily_hulls_weather, aes(x = Temperature, y = log(area_km2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Temperature (°C)",
    y = "log(Activity Space Area km²)",
    title = "Activity Space vs Temperature") +
  theme_minimal()
```

```{r}
# visualise the variable RainDur
ggplot(daily_hulls_weather, aes(x = RainDur, y = log(area_km2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Rain Duration (min)",
    y = "log(Activity Space Area km²)",
    title = "Activity Space vs Rain Duration") +
  theme_minimal()
```

```{r}
# linear model with log(area)
lm1 <- lm(log(area_km2) ~ Temperature + RainDur, data = daily_hulls_weather)
```

```{r}
# full model summary (R^2, adjusted R^2, p-values, residual std error)
summary(lm1)


par(mfrow = c(1, 1))
plot(lm1, which = 1)  # Residuals vs Fitted
plot(lm1, which = 2)  # Q-Q plot
plot(lm1, which = 3)  # Scale-Location
plot(lm1, which = 5)  # Residuals vs Leverage
```

The model is significant as a whole (F-statistic p = 0.028), meaning Temperature and RainDur together do explain some variation in activity space.  

There is a suggestive positive trend with Temperature, but we cannot confidently reject H0.  

The low R^2 (0.21) and adjusted R^2 (0.15) suggests other factors (e.g. day of week, work/leisure day, season) that influence activity space more than weather. 

## Influence of Weather on Walking

Travel Mode detection
Fokus nur auf on foot
--> calculate speed for each segment

--> assign weather information to each segment

# Results

# Discussion

# Appendix

### Wordcount

<!-- after installing the wordcountadding, remove the line "#| eval: false" -->

```{r}
#| eval: false
wordcountaddin::word_count("index.qmd")
```


