---
title: Weather Conditions and their Influence on human Mobility Patterns
subtitle: GEO 880 Project
author: Lars Weidinger & Javier Feller
date: today
format: 
  html:
    fig_caption: true
    fig-align: center
    highlight: tango
    number-sections: false
    code-fold: true
    theme: lumen
    embed-resources: true
    toc: true
    toc_depth: 2
    toc_float: true
execute:                      # Set global options for execution of code blocks
  echo: true                  # False = no code printed
  warning: false
  message: false
editor: source
lang: en
bibliography: bibliography.bib
---

```{r preprocessing}
#| code-summary: preprocessing
source("preprocessing.R") # this runs the whole preprocessing file
``` 

# Notes from Project Proposal Call
Meteo Daten Stunde vs. Tag --> ausprobieren

Sample zu klein wahrscheinlich, Stunde sieht man probably mehr Muster
andere Studierende? 

Map Matching --> nicht nötig für unser Projekt

Timeline vs. ArcGIS Earth --> wie kann man das vergleichen?
Filter machen für die Abschnitte, bei denen wir die gleichen Informationen haben
--> falls zu aufwändig, dann treffen wir eine pragmatische Entscheidung und halten das so fest (Vereinfachung) 

1) Anschauen, ob man mit Meteodaten (1 Tag) Unterschiede auf dem gesamten Datensatz sieht. 
- Zu Beginn explorativ anschauen, ob es Unterschiede gibt von Wochentagen und Wochenenden

2) Dann genauer schauen mit Meteodaten (1h oder 15min) --> Temperatur als Feld modellieren, nicht als Voronoi Polygone

3) Dann noch genauer schauen und zwar nur dort, wo man gelaufen ist.
- gibt es Unterschiede in Laufdistanz oder Laufgeschwindigkeit je nach Temperatur?

4) Falls man keine krassen Unterschiede sieht --> zwei Ansätze miteinander vergleichen (z.B. 15min vs. 1h oder 1 Tag) 
--> geht um konzeptueller Approach, Ergebnisse sind nicht so relevant

Alternative Datensätze: Tageslicht? Stürme? 

Ran Nathan's Movement Ecology Paradigm: wo befinden wir usn mit diesem Project, ist evtl. v.a. bei der Discussion relevant


# Abstract

# Introduction

@laube2011 investigate how temporal scale affects the calculation of movement parameters (speed, sinuosity and turning angle) of animal trajectories. 

# Material and Methods

## Load Packages
```{r}
library(pacman)

# Load and if not found in the library will attempt to install the package
pacman::p_load(here,
               sf,
               readr,
               tmap,
               dplyr,
               ggplot2,
               tidyr,
               jsonlite, # to import JSON files
               lubridate, # temporal import from JSON file
               tidyverse,
               leaflet, # interactive maps, more personal experience than on tmap
               tmap, # interactive maps
               concaveman, # construct concave hulls
               gstat, # for IDW
               stars, # to create a grid
               purrr,
               install = TRUE)

```

## Set CRS
```{r}
crs_lv95  <- 2056
crs_wgs84 <- 4326
```

## Import Data
See the preprocessing.R file where we import:  
- weather data  
- weather station location data  
- tracking data   
- city boundary data  

## Weather Data Daily (Voronoi Polygon)

### Locating Weather Stations
In the provided metadata JSON file the location of the weather stations can be found. We imported and cleaned the metadat in the preprocessing file. 

```{r}
ggplot() +
  geom_sf(data = stadtgrenze, fill = "#D5F4FF", color = "black") +
  geom_sf(data = weather_stations_sf, aes(color = ID), size = 3) +
  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Weather Stations within Zurich",
    color = "Station",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Constrcut Voronoi Polygons around Measuring Stations
Since we need to assign weather data to our stops and later our segments we need to relate the two datasets spatially. For that, we construct Voronoi polygons around the measuring stations and assign each tracking point the weather values from the station it is closest to (= inside the Voronoi polygon).

```{r}
# both datasets are in CH1903+

# create a union of points
pts_union <- st_union(weather_stations_sf)

# create Voronoi tessellation with an envelope
# st_voronoi has an envelope argument. Instead of generating a huge (or to small) tessellation we can set the scope here
envelope_stadtgrenze <- st_as_sfc(st_bbox(stadtgrenze))

voronoi <- st_voronoi(
  pts_union,
  envelope = envelope_stadtgrenze)

# extract polygons
voronoi_polygons <- st_collection_extract(voronoi, "POLYGON")

# convert to sf object
voronoi_sf <- st_sf(
  ID = weather_stations_sf$ID,
  geometry = st_cast(voronoi_polygons))

# clip to city boundary
voronoi_clipped <- sf::st_intersection(voronoi_sf, stadtgrenze)
```

```{r}
ggplot() +
  geom_sf(data = voronoi_clipped, aes(fill = ID)) +
  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Voronoi Polygons around\n Weather Stations in Zurich",
    color = "Station",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Assign Google Timeline Data to Voronoi Polygons
We can now use a spatial join to assign the corresponding weather station to each one of the Google Timeline tracking points.  
Since we traveled quite a lot outside of the city of Zurich, we omit the data outside the city (NA values), which is approximately 60% of the data points. 

```{r}
# both datasets are in CH1903+

# spatial join: assign polygon attributes (ID) to points
google_timeline_sf_lv95 <- st_join(
  google_timeline_sf_lv95,
  voronoi_clipped["ID"],
  join = st_within) |>
  
  drop_na() # omit NA values (quite a lot of them)
```

```{r}
# visualise timeline points
ggplot() +
  geom_sf(data = stadtgrenze, fill = "#D5F4FF", color = "black") +
  
  geom_sf(data = google_timeline_sf_lv95, color = "orange", size = 0.6) +

  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Google Timeline Data Points\n within Zurich",
    color = "Stop",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Assign Weather Data to Google Timeline Data
Now that we have assigned each data point to its closest weather station, we need to assign them their relevant weather values. We decided on using temperature (T) and rain (RainDur). One problem is, that the station Heubeeribüel does not collect RainDur information. 

```{r}
# based on the exercises of week 5, Task 3 ff we first create a join key so that we can later match them exactly

google_timeline_sf_lv95 <- google_timeline_sf_lv95 |>
  mutate(
    Date = as.Date(time)) # since we only have one measurement per night

weather_daily_df <- weather_daily_df |>
  mutate(
    Date = as.Date(Datum)) # since we only have one measurement per night
```

```{r}
# prepare weather data
weather_daily_df_wide <- weather_daily_df |>
  filter(Parameter %in% c("T", "RainDur")) |> # we only want T and RainDur
  select(Date, Standort, Parameter, Wert) |>
  pivot_wider(                               # we need the parameters as columns
    names_from = Parameter,
    values_from = Wert
  ) |>
  rename(
    Temperature = T,
    RainDur = RainDur, 
    ID = Standort
  )
```

```{r}
# join the two datasets
google_timeline_weather <- google_timeline_sf_lv95 |>
  left_join(
    weather_daily_df_wide,
    by = c("Date", "ID")
  )
```

```{r}
# visualise timeline points with voronoi polygons
ggplot() +
  geom_sf(data = voronoi_clipped, aes(fill = ID)) +
  
  geom_sf(data = google_timeline_sf_lv95, color = "orange", size = 0.6) +

  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Google Timeline Data Points\n within the Voronoi Polygons",
    color = "Stop",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

## Visualise Activity Space (Concave Hulls)

To answer RQ II we need to construct some sort of measurement for activity space. We decided on constructing concave hulls around the trips for each day. We then can calculate the area of these hulls, take the average weather values across the whole hull and analyse if the area is larger when T is higher or P is lower.

We decided to use concave hulls rather than convex hulls since they are less prone to outliers. Even though there are some convex hull implementations that allow for some sort of variable to be manipulated in order to exclude outliers (e.g. adehabitatHR::mcp()), the concave hull is more suited for our use case. We could have also used Alpha shapes (alphahull::) which accounts for holes in our data but this is not needed since we are dealing with cotinuous travel data. 

QUELLE: GEO 872, 3. Vlsg 02.10.24 Point Patterns - Polygon Delineation

Alternative: Standard devitational ellipse (SDE): forms a closed shape that represents distributional properties of the point set but does not enclose all points.

### Construct Concave Hulls

In the first approach we called concaveman() on the entire sfc column (all geometries at once), not on each row individually. It computed one hull over all points and that same result gets recycled into every row which resulted in all the hulls being the same. We asked Claude AI (Model Sonnet 4.6) for help and it delivered the following solution using lapply.

```{r}
# Step 1: union points by date
google_timeline_hulls <- google_timeline_sf_lv95 |>
  group_by(Date) |>
  summarise(geometry = st_union(geometry), .groups = "drop") |>
  st_as_sf()

# Step 2: build concave hull for each date
hulls_list <- lapply(google_timeline_hulls$geometry, function(g) {
  coords <- st_coordinates(g)[, c("X", "Y")]
  
  # need at least 3 unique points for a polygon
  if (is.null(nrow(coords)) || nrow(unique(coords)) < 3) return(NA)
  
  st_polygon(list(concaveman(coords)))
})

# Step 3: check which dates failed
failed_dates <- google_timeline_hulls$Date[sapply(hulls_list, function(x) identical(x, NA))]
print(failed_dates)  # inspect which dates are problematic

# Step 4: remove failed dates and assign geometry
google_timeline_hulls <- google_timeline_hulls[!sapply(hulls_list, function(x) identical(x, NA)), ]

hulls_list <- hulls_list[!sapply(hulls_list, function(x) identical(x, NA))]

# Step 5: assign hulls back as sfc column
google_timeline_hulls$geometry <- st_sfc(hulls_list, crs = st_crs(google_timeline_sf_lv95))
```

```{r}
# calculate area
google_timeline_hulls <- google_timeline_hulls |>
  mutate(
    area_m2 = as.numeric(st_area(geometry)),
    area_km2 = area_m2 / 1000000)
```

```{r}
# visualise concave hulls
ggplot() +
  geom_sf(data = stadtgrenze, fill = "transparent", color = "black") +
  
  geom_sf(data = google_timeline_hulls, aes(fill = as.factor(Date)), alpha = 0.8) +
  
  scale_fill_discrete(guide = "none") +
  
  labs(
    title = "Concave Hulls of Daily Movement in Zurich",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Assign Weather Values
```{r}
# average weather values across whole hull
daily_weather <- google_timeline_weather |>
  st_drop_geometry() |>
  group_by(Date) |>
  summarise(
    Temperature = mean(Temperature, na.rm = TRUE),
    RainDur = mean(RainDur, na.rm = TRUE)
  )

daily_hulls_weather <- google_timeline_hulls |>
  left_join(daily_weather, by = "Date")
```

For visualisation purpose we show four random concave hulls. As can be seen in the below visualisation, the temperature assignment worked. 
```{r}
# visualise concave hulls with weather data
ggplot() +
  geom_sf(data = stadtgrenze, fill = "transparent", color = "black") +
  
  geom_sf(data = daily_hulls_weather[c(1, 8, 24, 35), ], aes(fill = Temperature), alpha = 1) +
  
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Temperature [°C]") +
  
  labs(
    title = "Concave Hulls of Daily Movement in Zurich",
    color = "Temperature",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

```{r}
# visualise the variable Temperature
ggplot(daily_hulls_weather, aes(x = Temperature, y = log(area_m2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Temperature (°C)",
    y = "log(Activity Space Area km²)",
    title = "Activity Space vs Temperature") +
  theme_minimal()
```

```{r}
# visualise the variable RainDur
ggplot(daily_hulls_weather, aes(x = RainDur, y = log(area_m2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Rain Duration (min)",
    y = "log(Activity Space Area km²)",
    title = "Activity Space vs Rain Duration") +
  theme_minimal()
```

### Statistical Analysis
```{r}
# linear model with log(area)
lm1 <- lm(log(area_m2) ~ Temperature + RainDur, data = daily_hulls_weather)
```

```{r}
# full model summary (R^2, adjusted R^2, p-values, residual std error)
summary(lm1)

par(mfrow = c(1, 1))
plot(lm1, which = 1)  # Residuals vs Fitted
plot(lm1, which = 2)  # Q-Q plot
plot(lm1, which = 3)  # Scale-Location
plot(lm1, which = 5)  # Residuals vs Leverage
```

The model is significant as a whole (F-statistic p = 0.028), meaning Temperature and RainDur together do explain some variation in activity space.  

There is a suggestive positive trend with Temperature, but we cannot confidently reject H0.  

The low R^2 (0.21) and adjusted R^2 (0.15) suggests other factors (e.g. day of week, work/leisure day, season) that influence activity space more than weather. 

## Weather Daily (Field)
We only have weather measurements at at a handful of stations across the city. Assigning each location fix the weather data of the Voronoi polygon of the station it is in, is simplistic. Weather changes somewhat gradually across space, a more "fluid" approach would be better. That is why we decided on using Inverse Distance Weighting (IDW), which is a simple approach to generate gradually changing values across space using some sort of tessellation.  
With only few stations the interpolation accuracy will be low. We are not capturing real microclimate variation like urban heat island effects or cooling green spaces and water bodies. But since we plan on repeating our method with the finer scaled Meteoblue stations, we decided on applying the IDW method. 

Since we do not want to compute a full raster per day we will do a point-wise IDW prediction directly at each location fix's coordinates, using all weather station information. 

A raster per timestamp needs a lot of computation time and we only need the value at the fix's exact coordinates. Point-wise IDW is just a weighted average of all the distance, which takes much less time and we avoid avoid managing hundreds of rasters.

### Inverse Distance Weighting (IDW)

```{r}
# this code chunk was constructed with the help of Claude (Sonnet 4.6)

idw_for_date <- function(fixes_date, stations_date, power = 2) {
  # fixes_date: sf object for this date
  # stations_date: sf object with Temperature, RainDur for this date
  # power = 2 is the standard value
  
  # full distance matrix: n_fixes x n_stations
  dist_mat <- st_distance(fixes_date, stations_date)
  dist_mat <- matrix(as.numeric(dist_mat), nrow = nrow(fixes_date), ncol = nrow(stations_date))
  
  # avoid division by zero
  dist_mat[dist_mat == 0] <- 1e-9
  
  weights <- 1 / (dist_mat^power)
  weights_norm <- weights / rowSums(weights)
  
  fixes_date |>
    mutate(
      Temperature_IDW = as.numeric(weights_norm %*% stations_date$Temperature),

      RainDur_IDW = apply(weights_norm, 1,
                          function(w) {
                            valid <- !is.na(stations_date$RainDur) # ignore stations with missing RainDur
                            
                            if(sum(valid) == 0) {
                              return(NA)
                              }
                            
                            w_valid <- w[valid]
                            rain_valid <- stations_date$RainDur[valid]
                            
                            # renormalize weights only along valid stations
                            w_valid <- w_valid / sum(w_valid)
                            
                            sum(w_valid * rain_valid)
                            }
                          )
      )
}


# apply per date (groups are small: one per day x n_stations)
unique_dates <- unique(google_timeline_sf_lv95$Date)

result_list <- map(unique_dates, function(d) {
  fixes_date <- google_timeline_sf_lv95 |>
    filter(Date == d)
  stations_date <- weather_daily_df_wide |>
    filter(Date == d) |>
    inner_join(st_drop_geometry(weather_stations_sf) |>
                 select(ID), by = "ID") |>
    left_join(weather_stations_sf |>
                select(ID, geometry), by = "ID") |>
    st_as_sf()
  
  if (nrow(stations_date) == 0) {
    fixes_date$Temperature_IDW <- NA
    fixes_date$RainDur_IDW <- NA
    return(fixes_date)
  }
  
  idw_for_date(fixes_date, stations_date)
})

google_timeline_weather_daily <- bind_rows(result_list)
```

### Assign Weather Values
```{r}
# average weather values across whole hull
daily_weather_field <- google_timeline_weather_daily |>
  st_drop_geometry() |>
  group_by(Date) |>
  summarise(
    Temperature = mean(Temperature_IDW, na.rm = TRUE),
    RainDur = mean(RainDur_IDW, na.rm = TRUE)
  )

daily_hulls_weather_field <- google_timeline_hulls |>
  left_join(daily_weather_field, by = "Date")
```

```{r}
# visualise the variable Temperature
ggplot(daily_hulls_weather_field, aes(x = Temperature, y = log(area_m2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Temperature (°C)",
    y = "log(Activity Space Area km²)",
    title = "Activity Space vs Temperature") +
  theme_minimal()
```

```{r}
# visualise the variable RainDur
ggplot(daily_hulls_weather_field, aes(x = RainDur, y = log(area_m2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Rain Duration (min)",
    y = "log(Activity Space Area km²)",
    title = "Activity Space vs Rain Duration") +
  theme_minimal()
```

### Statistical Analysis
```{r}
# linear model with log(area)
lm2 <- lm(log(area_m2) ~ Temperature + RainDur, data = daily_hulls_weather_field)
```

```{r}
# full model summary (R^2, adjusted R^2, p-values, residual std error)
summary(lm2)
```

The P-value of the RainDur is slightly lower than from lm1, indicating that the interpolation of precipitation duration is a more appropriate way of identifiying a relationship between weather and movement.  
Furthermore, the R^2 and Adjusted R^2 values are slightly higher, meaning that our interpolated model explains a bit more of the values than the model with the Voronoi polygons. 

## Weather Hourly (Field)
Now that we have established that weather data is more accurate when using IDW to interpolate values between stations, we will repeat this process with the hourly weather data.  
The measurements are taken by the same stations as for the daily weather data. Therefore, we only really need to adapt the input data. The concave hulls will stay the same, we will still take all fixes from one day into account.

### Prepare Weather Data

```{r}
# prepare weather data
weather_hourly_df_wide <- weather_hourly_df |>
  filter(Parameter %in% c("T", "RainDur")) |> # we only want T and RainDur
  select(Date, Standort, Parameter, Wert) |>
  pivot_wider(                               # we need the parameters as columns
    names_from = Parameter,
    values_from = Wert
  ) |>
  rename(
    Temperature = T,
    RainDur = RainDur, 
    ID = Standort
  )
```



## Influence of Weather on Walking

Travel Mode detection
Fokus nur auf on foot
--> calculate speed for each segment

--> assign weather information to each segment

# Results

# Discussion

# Appendix

### Wordcount

<!-- after installing the wordcountadding, remove the line "#| eval: false" -->

```{r}
#| eval: false
wordcountaddin::word_count("index.qmd")
```


