---
title: Weather Conditions and their Influence on human Mobility Patterns
subtitle: GEO 880 Project
author: Lars Weidinger & Javier Feller
date: today
format: 
  html:
    fig_caption: true
    fig-align: center
    highlight: tango
    number-sections: false
    code-fold: true
    theme: lumen
    embed-resources: true
    toc: true
    toc_depth: 2
    toc_float: true
execute:                      # Set global options for execution of code blocks
  echo: true                  # False = no code printed
  warning: false
  message: false
editor: source
lang: en
bibliography: bibliography.bib
---

```{r preprocessing}
#| code-summary: preprocessing
source("preprocessing.R") # this runs the whole preprocessing file
``` 

# Notes from Project Proposal Call
Meteo Daten Stunde vs. Tag --> ausprobieren

Sample zu klein wahrscheinlich, Stunde sieht man probably mehr Muster
andere Studierende? 

Map Matching --> nicht nötig für unser Projekt

Timeline vs. ArcGIS Earth --> wie kann man das vergleichen?
Filter machen für die Abschnitte, bei denen wir die gleichen Informationen haben
--> falls zu aufwändig, dann treffen wir eine pragmatische Entscheidung und halten das so fest (Vereinfachung) 

1) Anschauen, ob man mit Meteodaten (1 Tag) Unterschiede auf dem gesamten Datensatz sieht. 
- Zu Beginn explorativ anschauen, ob es Unterschiede gibt von Wochentagen und Wochenenden

2) Dann genauer schauen mit Meteodaten (1h oder 15min) --> Temperatur als Feld modellieren, nicht als Voronoi Polygone


3) Dann noch genauer schauen und zwar nur dort, wo man gelaufen ist.
- gibt es Unterschiede in Laufdistanz oder Laufgeschwindigkeit je nach Temperatur?

4) Falls man keine krassen Unterschiede sieht --> zwei Ansätze miteinander vergleichen (z.B. 15min vs. 1h oder 1 Tag) 
--> geht um konzeptueller Approach, Ergebnisse sind nicht so relevant

Alternative Datensätze: Tageslicht? Stürme? 

Ran Nathan's Movement Ecology Paradigm: wo befinden wir usn mit diesem Project, ist evtl. v.a. bei der Discussion relevant


# Abstract

# Introduction

@laube2011 investigate how temporal scale affects the calculation of movement parameters (speed, sinuosity and turning angle) of animal trajectories. 

# Material and Methods

## Load Packages
```{r}
library(pacman)

# Load and if not found in the library will attempt to install the package
pacman::p_load(here,
               sf,
               readr,
               tmap,
               dplyr,
               ggplot2,
               tidyr,
               jsonlite, # to import JSON files
               lubridate, # temporal import from JSON file
               tidyverse,
               leaflet, # interactive maps, more personal experience than on tmap
               tmap, # interactive maps
               concaveman, # construct concave hulls
               gstat, # for IDW
               stars, # to create a grid
               install = TRUE)

```

## Set CRS
```{r}
crs_lv95  <- 2056
crs_wgs84 <- 4326
```

## Import Data
See the preprocessing.R file where we import:  
- weather data  
- weather station location data  
- tracking data   
- city boundary data  

## Visualise Weather Data Daily (Voronoi Polygon)

### Locating Weather Stations
In the provided metadata JSON file the location of the weather stations can be found. We imported and cleaned the metadat in the preprocessing file. 

```{r}
ggplot() +
  geom_sf(data = stadtgrenze, fill = "#D5F4FF", color = "black") +
  geom_sf(data = weather_stations_sf, aes(color = ID), size = 3) +
  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Weather Stations within Zurich",
    color = "Station",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Constrcut Voronoi Polygons around Measuring Stations
Since we need to assign weather data to our stops and later our segments we need to relate the two datasets spatially. For that, we construct Voronoi polygons around the measuring stations and assign each tracking point the weather values from the station it is closest to (= inside the Voronoi polygon).

```{r}
# both datasets are in CH1903+

# create a union of points
pts_union <- st_union(weather_stations_sf)

# create Voronoi tessellation with an envelope
# st_voronoi has an envelope argument. Instead of generating a huge (or to small) tessellation we can set the scope here
envelope_stadtgrenze <- st_as_sfc(st_bbox(stadtgrenze))

voronoi <- st_voronoi(
  pts_union,
  envelope = envelope_stadtgrenze)

# extract polygons
voronoi_polygons <- st_collection_extract(voronoi, "POLYGON")

# convert to sf object
voronoi_sf <- st_sf(
  ID = weather_stations_sf$ID,
  geometry = st_cast(voronoi_polygons))

# clip to city boundary
voronoi_clipped <- sf::st_intersection(voronoi_sf, stadtgrenze)
```

```{r}
ggplot() +
  geom_sf(data = voronoi_clipped, aes(fill = ID)) +
  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Voronoi Polygons around\n Weather Stations in Zurich",
    color = "Station",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Assign Google Timeline Data to Voronoi Polygons
We can now use a spatial join to assign the corresponding weather station to each one of the Google Timeline tracking points.  
Since we traveled quite a lot outside of the city of Zurich, we omit the data outside the city (NA values), which is approximately 60% of the data points. 

```{r}
# both datasets are in CH1903+

# spatial join: assign polygon attributes (ID) to points
google_timeline_sf_lv95 <- st_join(
  google_timeline_sf_lv95,
  voronoi_clipped["ID"],
  join = st_within) |>
  
  drop_na() # omit NA values (quite a lot of them)
```

```{r}
# visualise timeline points
ggplot() +
  geom_sf(data = stadtgrenze, fill = "#D5F4FF", color = "black") +
  
  geom_sf(data = google_timeline_sf_lv95, color = "orange", size = 0.6) +

  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Google Timeline Data Points\n within Zurich",
    color = "Stop",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Assign Weather Data to Google Timeline Data
Now that we have assigned each data point to its closest weather station, we need to assign them their relevant weather values. We decided on using temperature (T) and rain (RainDur). One problem is, that the station Heubeeribüel does not collect RainDur information. 

```{r}
# based on the exercises of week 5, Task 3 ff we first create a join key so that we can later match them exactly

google_timeline_sf_lv95 <- google_timeline_sf_lv95 |>
  mutate(
    Date = as.Date(time)) # since we only have one measurement per night

weather_daily_df <- weather_daily_df |>
  mutate(
    Date = as.Date(Datum)) # since we only have one measurement per night
```

```{r}
# prepare weather data
weather_daily_df_wide <- weather_daily_df |>
  filter(Parameter %in% c("T", "RainDur")) |> # we only want T and RainDur
  select(Date, Standort, Parameter, Wert) |>
  pivot_wider(                               # we need the parameters as columns
    names_from = Parameter,
    values_from = Wert
  ) |>
  rename(
    Temperature = T,
    RainDur = RainDur, 
    ID = Standort
  )
```

```{r}
# join the two datasets
google_timeline_weather <- google_timeline_sf_lv95 |>
  left_join(
    weather_daily_df_wide,
    by = c("Date", "ID")
  )
```

```{r}
# visualise timeline points with voronoi polygons
ggplot() +
  geom_sf(data = voronoi_clipped, aes(fill = ID)) +
  
  geom_sf(data = google_timeline_sf_lv95, color = "orange", size = 0.6) +

  
  coord_sf(datum = crs_lv95) +
  
  labs(
    title = "Google Timeline Data Points\n within the Voronoi Polygons",
    color = "Stop",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

## Visualise Activity Space (Concave Hulls)

To answer RQ II we need to construct some sort of measurement for activity space. We decided on constructing concave hulls around the trips for each day. We then can calculate the area of these hulls, take the average weather values across the whole hull and analyse if the area is larger when T is higher or P is lower.

We decided to use concave hulls rather than convex hulls since they are less prone to outliers. Even though there are some convex hull implementations that allow for some sort of variable to be manipulated in order to exclude outliers (e.g. adehabitatHR::mcp()), the concave hull is more suited for our use case. We could have also used Alpha shapes (alphahull::) which accounts for holes in our data but this is not needed since we are dealing with cotinuous travel data. 

QUELLE: GEO 872, 3. Vlsg 02.10.24 Point Patterns - Polygon Delineation

Alternative: Standard devitational ellipse (SDE): forms a closed shape that represents distributional properties of the point set but does not enclose all points.

### Construct Concave Hulls

In the first approach we called concaveman() on the entire sfc column (all geometries at once), not on each row individually. It computed one hull over all points and that same result gets recycled into every row which resulted in all the hulls being the same. We asked Claude AI (Model Sonnet 4.6) for help and it delivered the following solution using lapply.

```{r}
# Step 1: union points by date
google_timeline_hulls <- google_timeline_sf_lv95 |>
  group_by(Date) |>
  summarise(geometry = st_union(geometry), .groups = "drop") |>
  st_as_sf()

# Step 2: build concave hull for each date
hulls_list <- lapply(google_timeline_hulls$geometry, function(g) {
  coords <- st_coordinates(g)[, c("X", "Y")]
  
  # need at least 3 unique points for a polygon
  if (is.null(nrow(coords)) || nrow(unique(coords)) < 3) return(NA)
  
  st_polygon(list(concaveman(coords)))
})

# Step 3: check which dates failed
failed_dates <- google_timeline_hulls$Date[sapply(hulls_list, function(x) identical(x, NA))]
print(failed_dates)  # inspect which dates are problematic

# Step 4: remove failed dates and assign geometry
google_timeline_hulls <- google_timeline_hulls[!sapply(hulls_list, function(x) identical(x, NA)), ]

hulls_list <- hulls_list[!sapply(hulls_list, function(x) identical(x, NA))]

# Step 5: assign hulls back as sfc column
google_timeline_hulls$geometry <- st_sfc(hulls_list, crs = st_crs(google_timeline_sf_lv95))
```

```{r}
# calculate area
google_timeline_hulls <- google_timeline_hulls |>
  mutate(
    area_m2 = as.numeric(st_area(geometry)),
    area_km2 = area_m2 / 1000000)
```

```{r}
# visualise concave hulls
ggplot() +
  geom_sf(data = stadtgrenze, fill = "transparent", color = "black") +
  
  geom_sf(data = google_timeline_hulls, aes(fill = as.factor(Date)), alpha = 0.8) +
  
  scale_fill_discrete(guide = "none") +
  
  labs(
    title = "Concave Hulls of Daily Movement in Zurich",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

### Assign Weather Values
```{r}
# average weather values across whole hull
daily_weather <- google_timeline_weather |>
  st_drop_geometry() |>
  group_by(Date) |>
  summarise(
    Temperature = mean(Temperature, na.rm = TRUE),
    RainDur = mean(RainDur, na.rm = TRUE)
  )

daily_hulls_weather <- google_timeline_hulls |>
  left_join(daily_weather, by = "Date")
```

For visualisation purpose we show four random concave hulls. As can be seen in the below visualisation, the temperature assignment worked. 
```{r}
# visualise concave hulls with weather data
ggplot() +
  geom_sf(data = stadtgrenze, fill = "transparent", color = "black") +
  
  geom_sf(data = daily_hulls_weather[c(1, 8, 24, 35), ], aes(fill = Temperature), alpha = 1) +
  
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Temperature [°C]") +
  
  labs(
    title = "Concave Hulls of Daily Movement in Zurich",
    color = "Temperature",
    x = "Easting [m]",
    y = "Northing [m]") +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
    )
```

```{r}
# visualise the variable RainDur
ggplot(daily_hulls_weather, aes(x = RainDur, y = log(area_km2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Rain Duration (min)",
    y = "log(Activity Space Area km²)",
    title = "Activity Space vs Rain Duration") +
  theme_minimal()
```

```{r}
# linear model with log(area)
lm1 <- lm(log(area_km2) ~ Temperature + RainDur, data = daily_hulls_weather)
```

```{r}
# full model summary (R^2, adjusted R^2, p-values, residual std error)
summary(lm1)

par(mfrow = c(1, 1))
plot(lm1, which = 1)  # Residuals vs Fitted
plot(lm1, which = 2)  # Q-Q plot
plot(lm1, which = 3)  # Scale-Location
plot(lm1, which = 5)  # Residuals vs Leverage
```

The model is significant as a whole (F-statistic p = 0.028), meaning Temperature and RainDur together do explain some variation in activity space.  

There is a suggestive positive trend with Temperature, but we cannot confidently reject H0.  

The low R^2 (0.21) and adjusted R^2 (0.15) suggests other factors (e.g. day of week, work/leisure day, season) that influence activity space more than weather. 

## Visualise Weather Daily (Field)
We only have weather measurements at at a handful of stations across the city. Assigning each location fix the weather data of the Voronoi polygon of the station it is in, is simplistic. Weather changes somewhat gradually across space, a more "fluid" approach would be better. That is why we decided on using Inverse Distance Weighting (IDW), which is a simple approach to generate gradually changing values across space using some sort of tessellation.  
With only few stations the interpolation accuracy will be low. We are not capturing real microclimate variation like urban heat island effects or cooling green spaces and water bodies. But since we plan on repeating our method with the finer scaled Meteoblue stations, we decided on applying the IDW method. 

Since we do not want to compute a full raster per day we will do a point-wise IDW prediction directly at each location fix's coordinates, using all weather station information. 

A raster per timestamp needs a lot of computation time and we only need the value at the fix's exact coordinates. Point-wise IDW is just a weighted average of all the distance, which takes much less time and we avoid avoid managing hundreds of rasters.

### Inverse Distance Weighting (IDW)
```{r}
# this code chunk was constructed with the help of Claude 

# join station coordinates to daily weather data
weather_long <- weather_daily_df_wide |>
  left_join(st_drop_geometry(weather_stations_sf) |> 
              select(ID, geometry), by = c("Station_name" = "ID"))

# function: IDW prediction for one point given that day's station data
idw_predict <- function(x, y, date, weather_long, power = 2) {
  day_data <- weather_long |> filter(Date == date, !is.na(Temperature))
  
  if (nrow(day_data) == 0) return(NA_real_)
  
  dists <- sqrt((day_data$Koordinaten_LV95_X - x)^2 + (day_data$Koordinaten_LV95_Y - y)^2)
  
  # avoid division by zero if fix is exactly at a station
  if (any(dists == 0)) return(day_data$Temperature[which(dists == 0)[1]])
  
  weights <- 1 / (dists^power)
  sum(weights * day_data$Temperature) / sum(weights)
}

# apply to  Google timeline data (assumed columns: X, Y, Date)
timeline_data <- timeline_data |>
  rowwise() |>
  mutate(interpolated_temp = idw_predict(X, Y, Date, weather_long)) |>
  ungroup()
```



## Influence of Weather on Walking

```{r}
# Visualize ArcGisPro_trajectories
ggplot() +
  geom_sf(data = stadtgrenze, fill = "#D5F4FF", color = "black") +
  geom_sf(data = pts_lv95, aes(color = "Name"), size = 3) +
  
  coord_sf(crs = crs_lv95) +
  
  labs(
    title = "ArcGisPro trajectories in Zurich",
    color = "Station",
    x = "Easting [m]",
    y = "Northing [m]"
  ) +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(), 
    axis.title = element_blank()
  )
```

## Segmentation of trajectories (stops and moves)

```{r}
#### state function, define two timesteps using lead and nothing, selecting right column with data$columnname and apply function to later and now, which was defined before, and store it
difftime_secs <- function(later, now){
    as.numeric(difftime(later, now, units = "secs"))
}
```

```{r}
later <- lead(tracks_lv95$timestamp)
now <- tracks_lv95$timestamp

tracks_lv95$timelag <- difftime_secs(later, now)
```

```{r}
# sort after timestamp
tracks_lv95 <- tracks_lv95 |>
  arrange(timestamp)
```

```{r}
# compute distance and time difference, with support of Chatgpt
tracks_lv95 <- tracks_lv95 |>
  mutate(
    step_dist = as.numeric(st_distance(geometry, lead(geometry), by_element = TRUE)),
    step_time = as.numeric(difftime(lead(timestamp), timestamp, units = "secs"))
  )
```

```{r}
# define treshold of moving window, movement if distance of 1 step more than 5 
# meters in a time of less than 30 seconds.
tracks_lv95 <- tracks_lv95 |>
  mutate(
    moving_step = step_dist > 5 & step_time < 30
  )
```

```{r}
#assign new segment_id each time transition from stop to move or vice versa
#to connect the still separate points into segments, with support of Chatgpt
tracks_lv95 <- tracks_lv95 |>
  mutate(
    change = moving_step != lag(moving_step, default = first(moving_step)),
    segment_id = cumsum(change)
  )
```

```{r}
#calculate for each segment star time, end time, duration, geometry, speed
#with support of Chatgpt
segments <- tracks_lv95 |>
  group_by(segment_id, moving_step) |>
  summarise(
    start_time = first(timestamp),
    end_time = last(timestamp),

    duration = as.numeric(
      difftime(last(timestamp), first(timestamp), units = "secs")
    ),

    distance = sum(step_dist, na.rm = TRUE),

    speed = ifelse(duration > 0, distance / duration, NA_real_),

    speed_kmh = speed * 3.6,

    n_points = n(),
    geometry = st_union(geometry),
    .groups = "drop"
  )
```

```{r}
segments <- segments |>
  mutate(type = ifelse(moving_step, "MOVE", "STOP"))
```

```{r}
# Visualize stops and moves together with city boundary
ggplot() +
  geom_sf(data = stadtgrenze,
          fill = "#D5F4FF",
          color = "black") +
  geom_sf(data = segments,
          aes(color = type),
          size = 2) +
  scale_color_manual(
    values = c("STOP" = "blue", "MOVE" = "red")
  ) +
  coord_sf(crs = crs_lv95) +
  theme_minimal()
```


## speed distrubtion

```{r}
# distribution of speed to cut boundaries
quantile(segments$speed_kmh[segments$moving_step],
         probs=c(0.5,0.9,0.95, 0.98),
         na.rm = TRUE)
```

```{r}
#trim uper tail, with support of chatgpt
upper <- quantile(segments$speed_kmh, 0.975, na.rm = TRUE)

segments_trim <- segments |>
  filter(speed_kmh <= upper)
```

```{r}
# boxplot for distribution, with support of chatgpt
boxplot(
  segments_trim$speed_kmh,
  main = "Speed (km/h) – trimmed 2.5% tails",
  ylab = "Speed (km/h)",
  col = "lightgreen"
)
```
## Travel Mode detection (focus on walking)
```{r}
# create only walking segments (speed between 0 and 7 km/h)
walking_segments <- segments |> filter(segments$speed_kmh < 7 & segments$speed_kmh > 1 )
```


### Construct Voronoi Polygons around meteoblue measuring Stations to after/
### wards assign every segment to the closest meteoblue weather station
### this section with suppor of chatgpt

```{r}
# create geometry out of coords for meteoblue_stations
meteoblue_sf <- meteoblue_stations |> st_as_sf(coords=c("lonDecimal", "latDecimal"), crs=4326)
meteoblue_sf <- st_transform(meteoblue_sf, crs = crs_lv95)
```

```{r}
# create union of all station points and set city extent
pts_union <- st_union(st_geometry(meteoblue_sf))
envelope_zurich <- st_as_sfc(st_bbox(stadtgrenze))
```

```{r}
#create voronoi tesselation and extract polygons out of it, with support of Chatgpt
voronoi_meteoblue <- st_voronoi(pts_union, envelope=envelope_zurich)

voronoi_polygons <- st_collection_extract(
  voronoi_meteoblue,
  "POLYGON"
)
# convert to sf object
voronoi_sf <- st_sf(
  geometry = st_cast(voronoi_polygons, "POLYGON")
)
```

```{r}
# assign station attributes to polygons, with support of chatgpt
voronoi_sf$locationID <- meteoblue_sf$locationID

# clip to Zurich city boundary
voronoi_clipped_meteoblue <- st_intersection(
  voronoi_sf,
  stadtgrenze
)
```

```{r}
#join between voronoi_around stations and meteoblue weather data 
voronoi_weather <- voronoi_clipped_meteoblue |>
  left_join(
    meteoblue_weather_data,
    by = "locationID"
  )
```

```{r}
# visualize, with support of chatgpt
ggplot() +
  geom_sf(data = voronoi_clipped_meteoblue, aes(fill = locationID)) +
  geom_sf(data = stadtgrenze, fill = NA, color = "black") +
  geom_sf(data = meteoblue_sf, color = "red") +
  
  coord_sf(crs = crs_lv95) +
  
  scale_fill_viridis_d(option = "turbo") +
  guides(fill = "none") +
  
  labs(
    title = "Voronoi Polygons around meteoblue Stations in Zurich",
    fill = "Station",
    x = "Easting [m]",
    y = "Northing [m]"
  ) +
  
  theme_minimal() +
  
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text = element_blank(),
    axis.title = element_blank()
  )
```
```{r}
### Assign segments to voronoi-polygons, with support of chatgpt

#only look at point instead of segment
segment_points <- walking_segments |>
  st_cast("POINT")

#join with intersect
segment_points_weather <- st_join(
  segment_points,
  voronoi_clipped_meteoblue["locationID"],
  join = st_intersects
)

#aggregate back
meteoblue_segments <- segment_points_weather |>
  group_by(segment_id, locationID) |>
  summarise(geometry = st_union(geometry))

#clip to extent of Zurich
meteoblue_segments_clipped <- st_intersection(
  meteoblue_segments,
  stadtgrenze
)
```

```{r}
#visualize Movement Segments with meteoblue weather station polygons
ggplot() +
  geom_sf(data = voronoi_clipped_meteoblue, aes(fill = locationID), alpha = 0.3) +
  geom_sf(data = stadtgrenze, fill = NA, color = "black") +
  geom_sf(data = meteoblue_segments_clipped, color = "blue", size = 1) +
  
  coord_sf(crs = crs_lv95) +
  
  scale_fill_viridis_d(option = "turbo") +
  guides(fill = "none") +
  
  labs(
    title = "Movement Segments with meteoblue weather station polygons",
    x = "Easting",
    y = "Northing"
  ) +
  
  theme_minimal()
```

### temporal join of weather data and walking_segments
```{r}
# find most representative time (midtime) of each segment
segments <- segments |>
  mutate(
    mid_time = start_time + (end_time - start_time) / 2
  )
```

```{r}
meteoblue_weather_data <- meteoblue_weather_data |>
  mutate(timestamp = as.POSIXct(timestamp))
```

```{r}
# add spatial information to segments
segments_spatial <- st_join(
  segments,
  meteoblue_segments["locationID"],
  join = st_intersects
)
```

```{r}
#clean
segments_clean <- segments_spatial |>
  st_drop_geometry()
```

```{r}
segments_weather <- segments_clean |>
  left_join(meteoblue_weather_data, by = "locationID") |>
  mutate(
    time_diff = abs(as.numeric(difftime(mid_time, timestamp, units = "mins")))
  ) |>
  group_by(segment_id) |>
  slice_min(time_diff, n = 1, with_ties = FALSE) |>
  ungroup()
```

```{r}
#filter unrealistic values
segments_weather_clean <- segments_weather %>%
  filter(time_diff <= 120) %>%
  filter(
    speed_kmh >= quantile(speed_kmh, 0.02, na.rm = TRUE),
    speed_kmh <= quantile(speed_kmh, 0.98, na.rm = TRUE)
  )
```

```{r}
#check if temporal join worked
summary(segments_weather_clean$time_diff)
```
```{r}
#visualize
ggplot(segments_weather_clean, aes(x = value, y = speed_kmh)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  coord_cartesian(ylim = c(0, 10)) +
  theme_minimal() +
  labs(
    x = "Temperature (°C)",
    y = "Speed (km/h)"
  )
```

```{r}
trend <- lm(speed_kmh ~ value, data=segments_weather_clean)
summary(trend)
```

```{r}
segments_weather_clean <- segments_weather_clean %>%
  mutate(
    temp_group = cut(value,
                     breaks = quantile(value, probs = seq(0, 1, 0.25), na.rm = TRUE),
                     include.lowest = TRUE)
  )

lm(speed_kmh ~ temp_group, data = segments_weather_clean)
```
```{r}
anova(lm(speed_kmh ~ temp_group, data = segments_weather_clean))
```
```{r}
library(ggplot2)

ggplot(segments_weather_clean, aes(x = temp_group, y = speed_kmh)) +
  geom_boxplot(fill = "lightblue") +
  geom_jitter(width = 0.2, alpha = 0.3) +
  theme_minimal() +
  labs(
    x = "Temperature group",
    y = "Walking speed (km/h)"
  )
```
```{r}
ggplot(segments_weather_clean, aes(x = temp_group, y = speed_kmh)) +
  geom_boxplot(fill = "lightblue") +
  geom_jitter(width = 0.2, alpha = 0.3) +
  theme_minimal() +
  scale_x_discrete(labels = c(
    "(1,4.19]" = "Cold",
    "(4.2,9]" = "Cool",
    "(9,13.3]" = "Temperate",
    "(13.3,26.6]" = "Warm"
  )) +
  labs(
    x = "Temperature group",
    y = "Walking speed (km/h)"
  )
```
```{r}
segments_weather_clean <- segments_weather_clean %>%
  mutate(temp_group = factor(temp_group,
                             levels = levels(temp_group),
                             labels = c("Cold (1-4°C)", "Cool (4-9°C)", "Mild (9-13°C)", "Warm (13-26°C)")))
```

```{r}
segments_weather_clean <- segments_weather_clean %>%
  filter(!is.na(temp_group))
```

```{r}
ggplot(segments_weather_clean, aes(x = temp_group, y = speed_kmh)) +
  geom_boxplot(fill = "lightblue") +
  geom_jitter(width = 0.2, alpha = 0.3) +
  theme_minimal() +
  labs(
    x = "Temperature group",
    y = "Walking speed (km/h)"
  )
```

# analysis of acceleration-vs temperature



```{r}
sum(grepl("<gx:Track>", kml_text))
sum(grepl("</gx:Track>", kml_text))
```




# Results

# Discussion

# Appendix

### Wordcount

<!-- after installing the wordcountadding, remove the line "#| eval: false" -->

```{r}
#| eval: false
wordcountaddin::word_count("index.qmd")
```


