---
title: The title of your Semesterproject
subtitle: A subtle subtitle
author: Lars Weidinger & Javier Feller
format:
  html:
    code-fold: true
execute:
  warning: false
  message: false
lang: en
bibliography: bibliography.bib
---

```{r preprocessing}
#| code-summary: preprocessing
source("preprocessing.R") # this runs the whole preprocessing file
``` 

# Notes from Project Proposal Call
Meteo Daten Stunde vs. Tag --> ausprobieren

Sample zu klein wahrscheinlich, Stunde sieht man probably mehr Muster
andere Studierende? 

Map Matching --> nicht nötig für unser Projekt

Timeline vs. ArcGIS Earth --> wie kann man das vergleichen?
Filter machen für die Abschnitte, bei denen wir die gleichen Informationen haben
--> falls zu aufwändig, dann treffen wir eine pragmatische Entscheidung und halten das so fest (Vereinfachung) 

1) Anschauen, ob man mit Meteodaten (1 Tag) Unterschiede auf dem gesamten Datensatz sieht. 
- Zu Beginn explorativ anschauen, ob es Unterschiede gibt von Wochentagen und Wochenenden 

2) Dann genauer schauen mit Meteodaten (1h oder 15min) 

3) Dann noch genauer schauen und zwar nur dort, wo man gelaufen ist.
- gibt es Unterschiede in Laufdistanz oder Laufgeschwindigkeit je nach Temperatur?

4) Falls man keine krassen Unterschiede sieht --> zwei Ansätze miteinander vergleichen (z.B. 15min vs. 1h oder 1 Tag) 
--> geht um konzeptueller Approach, Ergebnisse sind nicht so relevant

Alternative Datensätze: Tageslicht? Stürme? 


# Abstract

# Introduction

@laube2011 investigate how temporal scale affects the calculation of movement parameters (speed, sinuosity and turning angle) of animal trajectories. 

# Material and Methods

## Load Packages
```{r}
library(pacman)

# Load and if not found in the library will attempt to install the package
pacman::p_load(here,
               sf,
               readr,
               tmap,
               dplyr,
               ggplot2,
               tidyr,
               jsonlite,
               lubridate, # temporal import from JSON file
               tidyverse,
               leaflet, # interactive maps
               tmap, # interactive maps
               install = TRUE)

```

## Set CRS
```{r}
crs_lv95  <- 2056
crs_wgs84 <- 4326
```


## Import Data
See the preprocessing.R file where we import:  
- weather data  
- tracking data  
- boundary data?

```{r}
stadtgrenze <- read_sf("data/Gemeindegrenzen_OGD/Gemeindegrenzen_-OGD.gpkg", layer = "up_gemeinden_f") |> # take the polygon
  filter(gemeindename == "Zürich") |> # of the city of Zurich
  select(gemeindename, geom)
```

## Visualise Weather Data Daily

### Locating Weather Stations
Since the data did not provide coordinates of the measuring stations, we manually took the CH1903+ coordinates of points that have to be close to the actual measuring stations using map.geo.admin.ch. We will ask the City of Zurich for the exact locations of the stations but for now this is our approximation for the moment. 

```{r}
weather_stations <- data.frame(
  Standort = c(
    "Zch_Heubeeribüel",
    "Zch_Rosengartenstrasse",
    "Zch_Schimmelstrasse",
    "Zch_Stampfenbachstrasse"
  ),
  E = c(
    2685280.12,
    2682050.59,
    2682105.79,
    2683454.71
  ),
  N = c(
    1248962.89,
    1249880.93,
    1247104.63,
    1248019.62
  )
)

weather_stations_sf <- st_as_sf(
  weather_stations,
  coords = c("E", "N"),
  crs = crs_lv95)
```

### Constrcut Voronoi Polygons around Measuring Stations
Since we need to assign weather data to our stops and later our segments we need to relate the two datasets spatially. For that, we construct Voronoi polygons around the measuring stations and assign each tracking point the weather values from the station it is closest to (= inside the Voronoi polygon).

```{r}
# both datasets are in CH1903+

# create a union of points
pts_union <- st_union(weather_stations_sf)

# create Voronoi tessellation
voronoi <- st_voronoi(pts_union)

# extract polygons
voronoi_polygons <- st_collection_extract(voronoi, "POLYGON")

# convert to sf object
voronoi_sf <- st_sf(
  Standort = weather_stations_sf$Standort,
  geometry = st_cast(voronoi_polygons))

# clip to city boundary
voronoi_clipped <- st_intersection(voronoi_sf, stadtgrenze)
```

```{r}
plot(voronoi_clipped) 
```

### Assign Google Timeline Data to Voronoi Polygons
We can now use a spatial join to assign the corresponding weather station to each one of the Google Timeline tracking points.  
Since we traveled quite a lot outside of the city of Zurich, we omit the data outside the city (NA values), which is approximately 60% of the data points. 

```{r}
# both datasets are in CH1903+

# spatial join: assign polygon attributes (Standort) to points
google_timeline_sf_lv95 <- st_join(
  google_timeline_sf_lv95,
  voronoi_clipped["Standort"],
  join = st_within) |>
  
  drop_na() # omit NA values (quite a lot of them)
```

### Assign Weather Data to Google Timeline Data
Now that we have assigned each data point to its closest weather station, we need to assign them their relevant weather values. We decided on using temperature (T) and rain (RainDur).

```{r}

```

Zeitkomponente noch reinbringen bei der Analyse. 

## Visualise Trip Data

Concave Hull around Trips per Day 
--> calculate area

Average weather values across whole Hull

Compare if area is bigger when T is higher or P lower


## Influence of Weather on Walking
Travel Mode detection
Fokus nur auf on foot
--> calculate speed for each segment

--> assign weather information to each segment


# Results

# Discussion

# Appendix

### Wordcount

<!-- after installing the wordcountadding, remove the line "#| eval: false" -->

```{r}
#| eval: false
wordcountaddin::word_count("index.qmd")
```


